Assumes that data types, variables, etc have already been introduced. 3 minutes is not enough time to introduce scope, pass by copy, etc, that would be more material later.

# Personal Functions

We've already introduced the idea of putting a series of commands together. Wouldn't it be great to group a series of data transformations together into functions that we could re-use later?

We've already been using `R`'s built in functions in our analyses, functions such as `min`, `max`, `std`, `plot`. Similar to other languages, `R` also makes it fairly easy to write our own functions.

## Basic Function

`R` has a function called `function` to make **new** functions. Just as we previously assigned `values` to `variables` (i.e. `myVar <- 1`), we do a similar process for `functions`:

```{r}
myFunction <- function(){

}
```

Here we are assigning a `function` definition to `myFunction`. This particular function is not very useful, as it does not take any arguments, do anything, or return any values. But it is a valid function that can be called from the command line:

```{r}
myFunction()
```

Notice that this returns a `NULL` value, or **nothing** in the language of `R`.

## Return a Value

Lets change this so we actually assign and return a value:

```{r}
myFunction <- function(){
  x <- 2
  return(x)
}

```

Notice all we have done is write our usual variable assignment in the function body. Then, when we call the function, the value of the variable (`x` in this case) is returned.

```{r}
myFunction()
```

We can call this like any other function, including assigning the value that is returned to a new variable:

```{r}
myVar <- myFunction()
myVar
```

## Input Variables

Now let's make the function do something useful with input varibles. We can start by simply having a function square the value of x.

```{r}
myFunction <- function(x){
  x2 <- x ^ 2
  return(x2)
}
```

We can use this function like any other function in R.

```{r}
myFunction(x = 2)

v <- c(1, 2, 3, 4, 5)
myFunction(v)
```

In pairs or groups of 3, think about the functions you know in R and write your own custom function that takes two input values (x & y) and adds them together. Can you give y a default value of 0, so that if y is not entered the value of x is returned?





